---
title: "MOUS Group Report"
subtitle: "Cohort-level MEG pipeline summary"
date: last-modified
params:
  summary_json: "derivatives/mous_pipeline/group_summary.json"
---

```{python}
import json
from pathlib import Path
import pandas as pd
import numpy as np

summary_json = Path(r.params["summary_json"])
summary = json.loads(summary_json.read_text())
derivatives_root = summary_json.parent
rows = []
for metric, stats in summary.get("metrics", {}).items():
    rec = {"metric": metric}
    rec.update(stats)
    rows.append(rec)
tbl = pd.DataFrame(rows)

manifest_rows = []
for mf in sorted(derivatives_root.glob("*/m9_orchestration/*_run_manifest.json")):
    try:
        payload = json.loads(mf.read_text())
        metrics = payload.get("metrics", {}) or {}
    except Exception:
        continue
    sid = mf.parent.parent.name.removeprefix("sub-")
    manifest_rows.append(
        {
            "subject_id": sid,
            "n_trials_zinnen": metrics.get("n_zinnen"),
            "n_trials_woorden": metrics.get("n_woorden"),
            "mean_FD": metrics.get("m10_mean_fd"),
            "n_motion_outliers": metrics.get("m10_n_motion_outliers"),
            "ICA_components_removed": metrics.get("m2_ica_n_components_removed"),
            "DCI_zinnen": metrics.get("dci_zinnen"),
            "DCI_rest": metrics.get("dci_rest"),
            "p_task_vs_rest": metrics.get("p_task_vs_rest"),
            "prestim_beta_t": metrics.get("aim1_prestim_auc"),
            "N400m_r": metrics.get("aim1_n400m_zinnen_vs_woorden_t"),
            "MTG_spearman_r": (
                (metrics.get("m11_coupling") or {}).get("zinnen_prestim_beta_vs_mtg", {}) or {}
            ).get("r")
            if isinstance(metrics.get("m11_coupling"), dict)
            else None,
        }
    )
qc_tbl = pd.DataFrame(manifest_rows)

group_trials_path = derivatives_root / "group_trials.csv"
group_trials = pd.read_csv(group_trials_path) if group_trials_path.exists() else pd.DataFrame()
```

```{r}
library(ggplot2)
library(jsonlite)
library(knitr)
library(readr)

summary_json <- params$summary_json
derivatives_root <- dirname(summary_json)
summary_r <- jsonlite::fromJSON(summary_json, simplifyVector = FALSE)
metrics_r <- summary_r$metrics
if (is.null(metrics_r)) {
  metrics_r <- list()
}

metric_rows <- lapply(names(metrics_r), function(metric) {
  stats <- metrics_r[[metric]]
  data.frame(
    metric = metric,
    mean = if (!is.null(stats$mean)) stats$mean else NA_real_,
    sd = if (!is.null(stats$sd)) stats$sd else if (!is.null(stats$std)) stats$std else NA_real_,
    stringsAsFactors = FALSE
  )
})
tbl <- if (length(metric_rows) > 0) {
  do.call(rbind, metric_rows)
} else {
  data.frame(metric = character(), mean = numeric(), sd = numeric())
}

group_trials_path <- file.path(derivatives_root, "group_trials.csv")
group_trials <- if (file.exists(group_trials_path)) {
  readr::read_csv(group_trials_path, show_col_types = FALSE)
} else {
  data.frame()
}

theme_mous <- function(base_size = 13) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title       = element_text(face = "bold", size = base_size + 1, color = "#0f172a"),
      plot.subtitle    = element_text(color = "#64748b", size = base_size - 1, margin = margin(b = 8)),
      axis.title       = element_text(color = "#475569", size = base_size - 1),
      axis.text        = element_text(color = "#64748b"),
      panel.grid.major = element_line(color = "#e2e8f0"),
      panel.grid.minor = element_blank()
    )
}
```

## Group Summary Table

```{python}
tbl.style.format(precision=4)
```

## Integrated Aims Narrative

This group report presents the three aims as a coherent chain: trial-level preparation effects (Aim 1), multimodal MTG coupling (Aim 2), and traveling-wave directionality constraints (Aim 3). Null outcomes are interpreted as boundary conditions, not failures.

## Aim 1 - Group Readiness Signals

```{python}
aim1_metrics = [
    "dci_zinnen",
    "p_zinnen_vs_woorden",
]
if not tbl.empty:
    aim1_tbl = tbl[tbl["metric"].isin(aim1_metrics)].copy()
    aim1_tbl.style.format(precision=4) if not aim1_tbl.empty else print("No Aim 1 aggregate rows found.")
else:
    print("No group metric table available.")
```

```{r}
#| label: fig-aim1-group-prestim
#| fig-cap: "Group trial-level prestim beta by condition pooled across subjects."
#| fig-width: 8
#| fig-height: 4.5

if (nrow(group_trials) > 0 && all(c("prestim_beta", "condition") %in% names(group_trials))) {
  group_trials$condition <- factor(group_trials$condition)
  ggplot(group_trials, aes(x = condition, y = prestim_beta, fill = condition)) +
    geom_violin(trim = FALSE, alpha = 0.4, color = "white", linewidth = 0.2) +
    geom_boxplot(width = 0.12, outlier.shape = NA, alpha = 0.8, linewidth = 0.6) +
    scale_fill_manual(values = c("ZINNEN" = "#2563eb", "WOORDEN" = "#16a34a", "REST" = "#d97706")) +
    theme_mous() +
    theme(legend.position = "none") +
    labs(title = "Aim 1 cohort prestim beta contrast", x = NULL, y = "Pre-stim beta (a.u.)")
}
```

## Aim 2 - Group MEG-fMRI Coupling

```{python}
m11 = summary.get("metrics", {}).get("m11_coupling", {})
rows = []
for key in ("zinnen_prestim_beta_vs_mtg", "woorden_prestim_beta_vs_mtg"):
    stats = m11.get(key)
    if isinstance(stats, dict):
        rows.append(
            {
                "contrast": key,
                "mean_r": stats.get("mean"),
                "median_r": stats.get("median"),
                "wilcoxon_p": stats.get("wilcoxon_p"),
                "n_subjects": stats.get("n"),
            }
        )
pd.DataFrame(rows).style.format(precision=4) if rows else print("No Aim 2 coupling summaries found.")
```

## Aim 3 - Group Traveling-Wave Constraints

```{python}
aim3_rows = []
for key in ("dci_zinnen", "dci_rest", "p_task_vs_rest", "m12_wave_validation_z"):
    stats = summary.get("metrics", {}).get(key)
    if isinstance(stats, dict):
        aim3_rows.append({"metric": key, "mean": stats.get("mean"), "median": stats.get("median"), "p": stats.get("wilcoxon_p")})
pd.DataFrame(aim3_rows).style.format(precision=4) if aim3_rows else print("No Aim 3 summary rows found.")
```

## Group Means by Metric

```{r}
#| label: fig-group-means
#| fig-cap: "Group-level mean ± SD for each pipeline metric, sorted by mean value."
#| fig-width: 9
#| fig-height: 5

df <- as.data.frame(tbl)
if ("mean" %in% names(df)) {
  df <- df[order(df$mean), ]
  df$metric <- factor(df$metric, levels = df$metric)

  p <- ggplot(df, aes(x = metric, y = mean, fill = mean)) +
    geom_col(alpha = 0.85, width = 0.7) +
    scale_fill_gradient(low = "#bfdbfe", high = "#1d4ed8") +
    theme_mous() +
    theme(legend.position = "none") +
    coord_flip() +
    labs(
      title    = "Group mean by metric",
      subtitle = "Sorted ascending; colour intensity encodes magnitude",
      x = NULL, y = "Mean"
    )

  if ("sd" %in% names(df)) {
    p <- p + geom_errorbar(
      aes(ymin = mean - sd, ymax = mean + sd),
      width = 0.25, color = "#475569", linewidth = 0.6
    )
  }

  print(p)
}
```

## Subject-Level QC Transparency Table

```{python}
if not qc_tbl.empty:
    qc_tbl = qc_tbl.sort_values("subject_id").reset_index(drop=True)
    qc_tbl.style.format(precision=4)
else:
    print("No manifest-derived QC rows available.")
```

## m11 Primary vs Exploratory and FDR

```{python}
m11 = summary.get("metrics", {}).get("m11_coupling", {})
rows = []
for key, stats in m11.items():
    if not (isinstance(key, str) and key.endswith("_vs_mtg") and isinstance(stats, dict)):
        continue
    rows.append(
        {
            "contrast": key,
            "mean_r": stats.get("mean"),
            "wilcoxon_p": stats.get("wilcoxon_p"),
            "n_subjects": stats.get("n"),
            "feature_role": "primary" if "prestim_beta" in key else "exploratory",
        }
    )
if rows:
    pd.DataFrame(rows).sort_values(["feature_role", "contrast"]).style.format(precision=4)
else:
    print("No m11 coupling summaries present in group summary.")
```
