---
title: "MOUS Style Preview"
subtitle: "Placeholder content for visual design review"
date: last-modified
---

::: {.callout-note}
## Preview only
This page uses placeholder text and synthetic values so you can evaluate layout, typography, tables, and chart styling.
:::

## Executive Snapshot

```{r}
library(knitr)

summary_tbl <- data.frame(
  Metric = c(
    "Pipeline status",
    "Subjects processed",
    "Mean DCI (task)",
    "Rayleigh p",
    "Model AUC"
  ),
  Value = c("READY", "24", "0.63", "0.0042", "0.78")
)

kable(summary_tbl, caption = "Synthetic report summary values for style preview")
```

## Placeholder Figure

```{r}
#| label: fig-placeholder-timeseries
#| fig-cap: "Synthetic time-series preview for style checking."
#| fig-width: 9
#| fig-height: 3.5

library(ggplot2)
set.seed(42)

t <- seq(0, 8, length.out = 200)
y <- 0.52 + 0.08 * sin(2 * pi * t / 3) + rnorm(length(t), sd = 0.015)
df <- data.frame(time_s = t, dci = y)

ggplot(df, aes(time_s, dci)) +
  geom_ribbon(aes(ymin = 0.5, ymax = pmax(dci, 0.5)), fill = "#bfdbfe", alpha = 0.6) +
  geom_line(color = "#2563eb", linewidth = 1.1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "#94a3b8") +
  theme_minimal(base_size = 13) +
  labs(title = "Synthetic DCI over time", x = "Time (s)", y = "DCI")
```

## Placeholder Category Plot

```{r}
#| label: fig-placeholder-bars
#| fig-cap: "Synthetic category means with error bars."
#| fig-width: 8
#| fig-height: 4.5

library(ggplot2)

cats <- data.frame(
  condition = c("ZINNEN", "WOORDEN", "REST"),
  mean = c(0.68, 0.61, 0.49),
  sd = c(0.06, 0.05, 0.04)
)

ggplot(cats, aes(condition, mean, fill = condition)) +
  geom_col(alpha = 0.85, width = 0.65) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.18) +
  scale_fill_manual(values = c("ZINNEN" = "#2563eb", "WOORDEN" = "#16a34a", "REST" = "#d97706")) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none") +
  labs(title = "Synthetic condition means", x = NULL, y = "Value")
```

## Placeholder Notes

- Body text rhythm and heading hierarchy
- TOC card style and left navigation behavior
- Table striping + hover behavior
- Figure captions and code-fold controls

```{r}
# Quick code chunk so you can see code-fold UI in action.
preview_status <- function() {
  "Design preview active"
}

preview_status()
```
