---
title: "MOUS Subject Report"
subtitle: "Cumulative single-subject MEG pipeline summary"
date: last-modified
params:
  subject: "A2002"
  export_dir: "derivatives/mous_pipeline/sub-A2002/m8_reports/exports"
---

```{python}
import json
from pathlib import Path
import pandas as pd
from IPython.display import HTML

subject = params["subject"]
export_dir = Path(params["export_dir"])
metrics = json.loads((export_dir / f"{subject}_metrics.json").read_text())

directions_path = export_dir / f"{subject}_directions.csv"
sliding_path = export_dir / f"{subject}_sliding_dci.csv"
source_dirs_path = export_dir / f"{subject}_source_directions.csv"
source_dci_path = export_dir / f"{subject}_source_dci.csv"
source_roi_path = export_dir / f"{subject}_source_roi_summary.csv"
trials_path = export_dir / f"{subject}_trials.csv"
joined_path = export_dir / f"{subject}_trials_joined.csv"

directions = pd.read_csv(directions_path) if directions_path.exists() else None
sliding = pd.read_csv(sliding_path) if sliding_path.exists() else None
source_dirs = pd.read_csv(source_dirs_path) if source_dirs_path.exists() else None
source_dci = pd.read_csv(source_dci_path) if source_dci_path.exists() else None
source_roi = pd.read_csv(source_roi_path) if source_roi_path.exists() else None
trials = pd.read_csv(trials_path) if trials_path.exists() else None
joined = pd.read_csv(joined_path) if joined_path.exists() else None
```

```{r}
library(ggplot2)
library(dplyr)
library(knitr)
library(lmerTest)
library(readr)

theme_mous <- function(base_size = 13) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title       = element_text(face = "bold", size = base_size + 1, color = "#0f172a"),
      plot.subtitle    = element_text(color = "#64748b", size = base_size - 1, margin = margin(b = 8)),
      axis.title       = element_text(color = "#475569", size = base_size - 1),
      axis.text        = element_text(color = "#64748b"),
      panel.grid.major = element_line(color = "#e2e8f0"),
      panel.grid.minor = element_blank(),
      strip.text       = element_text(face = "bold", color = "#1e293b"),
      legend.text      = element_text(size = 10, color = "#475569")
    )
}

COND_COLORS <- c("ZINNEN" = "#2563eb", "WOORDEN" = "#16a34a", "REST" = "#d97706")
subject <- params$subject
export_dir <- params$export_dir
```

# Pilot and Pipeline Summary

## Run Snapshot

```{python}
HTML('<div style="color:#475569;">Decision status presentation removed from report output.</div>')
```

## Key Metrics

```{python}
_labels = {
    "run_status": "Run status",
    "n_trials": "Total trials",
    "n_zinnen": "ZINNEN trials",
    "n_woorden": "WOORDEN trials",
    "n_rest": "REST pseudo-epochs",
    "dci_zinnen": "Mean DCI - ZINNEN",
    "dci_woorden": "Mean DCI - WOORDEN",
    "dci_rest": "Mean DCI - REST",
    "p_task_vs_rest": "p (task vs REST, permutation)",
    "p_woorden_vs_rest": "p (WOORDEN vs REST)",
    "p_zinnen_vs_woorden": "p (ZINNEN vs WOORDEN)",
    "p_rayleigh_zinnen": "Rayleigh p - ZINNEN",
    "p_rayleigh_woorden": "Rayleigh p - WOORDEN",
    "aim1_prestim_auc": "Pre-stim beta AUC (Aim 1)",
    "aim1_prestim_auc_guard_reason": "Pre-stim AUC guard reason (Aim 1)",
    "aim1_prestim_is_degenerate": "Pre-stim degenerate flag (Aim 1)",
    "aim1_prestim_std": "Pre-stim finite std (Aim 1)",
    "aim1_prestim_n_unique": "Pre-stim unique finite count (Aim 1)",
    "aim1_prestim_n_finite": "Pre-stim finite count (Aim 1)",
    "aim1_n400m_zinnen_vs_woorden_t": "N400m t (Aim 1)",
    "aim1_dci_block_effect_p": "DCI block-effect p (Aim 1)",
    "aim1_dci_first_trial_effect_p": "DCI first-trial effect p (Aim 1)",
    "aim1_prestim_block_effect_p": "Pre-stim block-effect p (Aim 1)",
    "aim1_prestim_first_trial_effect_p": "Pre-stim first-trial effect p (Aim 1)",
    "aim1_prestim_distribution_artifact": "Pre-stim distribution artifact",
    "aim1_prestim_topography_artifact": "Pre-stim topography artifact",
    "aim1_prestim_condition_contrast_artifact": "Pre-stim condition contrast artifact",
    "aim1_n400m_spec": "N400m extraction spec",
    "alpha_dci_zinnen": "Alpha DCI - ZINNEN",
    "alpha_dci_woorden": "Alpha DCI - WOORDEN",
    "alpha_dci_rest": "Alpha DCI - REST",
    "alpha_dci_status": "Alpha DCI status",
    "m10_hrf_lag_sweep_s": "HRF lag sweep values (s)",
    "m10_hrf_lag_sweep_tables": "HRF lag sweep tables",
    "m11_hrf_lag_sweep_ready": "HRF lag sweep-ready labels",
    "m11_coupling_hrf_lag_sweep": "m11 HRF lag sweep coupling",
    "m10_hrf_lag_selection": "HRF lag selection provenance",
    "aim1_source_spatial_summary": "Source-space spatial summary",
    "analysis_decisions_artifact": "Analysis decisions artifact",
    "aim3_two_dipole_z_threshold": "Aim3 z-threshold",
    "aim3_two_dipole_passes_threshold": "Aim3 passes z-threshold",
    "aim3_two_dipole_z_margin": "Aim3 z-threshold margin",
    "m12_decision": "m12 threshold decision",
    "m10_n_trials_joined": "Joined MEG-fMRI trials",
}

def _fmt(v):
    if isinstance(v, float):
        return f"{v:.4g}"
    return str(v)

rows = [
    {"Metric": _labels.get(k, k), "Value": _fmt(metrics[k])}
    for k in _labels
    if k in metrics
]
pd.DataFrame(rows).style.hide(axis="index")
```

## Compatibility and Diagnostics

```{python}
diag_keys = [
    "aim1_prestim_auc_guard_reason",
    "aim1_prestim_distribution_artifact",
    "aim1_prestim_topography_artifact",
    "m10_hrf_lag_sweep_tables",
    "m11_coupling_hrf_lag_sweep",
    "m12_null_summary",
    "m12_decision",
    "analysis_decisions_artifact",
]
diag_rows = [{"Key": k, "Value": _fmt(metrics[k])} for k in diag_keys if k in metrics]
if diag_rows:
    pd.DataFrame(diag_rows).style.hide(axis="index")
else:
    print("No optional compatibility diagnostics present.")
```

## Wave Direction Distributions

```{r}
#| label: fig-directions
#| fig-cap: "Polar histograms of wave propagation directions per condition (36 bins)."
#| fig-width: 10
#| fig-height: 4

dirs_file <- file.path(export_dir, paste0(subject, "_directions.csv"))
if (file.exists(dirs_file)) {
  dirs <- read_csv(dirs_file, show_col_types = FALSE)
  ggplot(dirs, aes(x = direction_rad, fill = condition)) +
    geom_histogram(bins = 36, position = "identity", alpha = 0.75, color = "white", linewidth = 0.2) +
    coord_polar() +
    facet_wrap(~condition, nrow = 1) +
    scale_fill_manual(values = COND_COLORS) +
    theme_mous() +
    theme(legend.position = "none") +
    labs(
      title = paste("Wave direction distributions -", subject),
      subtitle = "Each panel shows the angular distribution of sensor-space traveling waves",
      x = NULL, y = "Count"
    )
} else {
  cat("No direction CSV found.\n")
}
```

## Sliding DCI

```{r}
#| label: fig-sliding-dci
#| fig-cap: "Directionality Consistency Index (DCI) over time for ZINNEN trials. Dashed line = chance (0.5)."
#| fig-width: 9
#| fig-height: 3.5

slide_file <- file.path(export_dir, paste0(subject, "_sliding_dci.csv"))
if (file.exists(slide_file)) {
  slide <- read_csv(slide_file, show_col_types = FALSE)
  ggplot(slide, aes(x = time_s, y = dci)) +
    geom_ribbon(aes(ymin = 0.5, ymax = pmax(dci, 0.5)), fill = "#bfdbfe", alpha = 0.5) +
    geom_line(color = "#2563eb", linewidth = 1.1) +
    geom_hline(yintercept = 0.5, linetype = "dashed", color = "#94a3b8", linewidth = 0.7) +
    theme_mous() +
    labs(
      title = paste("Sliding DCI - ZINNEN:", subject),
      subtitle = "Blue shading = above-chance directionality",
      x = "Time (s)", y = "DCI"
    )
} else {
  cat("No sliding DCI CSV found.\n")
}
```

# Aim 1 - Trial-level Readiness

```{r}
#| code-fold: true

trials_file <- file.path(export_dir, paste0(subject, "_trials.csv"))
if (file.exists(trials_file)) {
  trials <- read_csv(trials_file, show_col_types = FALSE)
  req_cols <- c("prestim_beta", "condition", "pos_in_block", "block_id")
  if (all(req_cols %in% names(trials))) {
    trials$condition <- factor(trials$condition)
    fit <- lmer(prestim_beta ~ condition + pos_in_block + (1 | block_id), data = trials)
    summary(fit)
  } else {
    missing_cols <- setdiff(req_cols, names(trials))
    knitr::asis_output(
      paste0(
        "Aim 1 model skipped (missing trial columns): ",
        paste(missing_cols, collapse = ", ")
      )
    )
  }
} else {
  knitr::asis_output("Trials CSV not available; Aim 1 model skipped.")
}
```

```{r}
#| label: fig-prestim-beta
#| fig-cap: "Distribution of pre-stimulus beta power across conditions."
#| fig-width: 7
#| fig-height: 4.5

if (file.exists(trials_file)) {
  trials <- read_csv(trials_file, show_col_types = FALSE)
  req_cols <- c("prestim_beta", "condition")
  if (all(req_cols %in% names(trials))) {
    trials$condition <- factor(trials$condition)
    ggplot(trials, aes(x = condition, y = prestim_beta, fill = condition, color = condition)) +
      geom_violin(trim = FALSE, alpha = 0.35, linewidth = 0.6) +
      geom_boxplot(width = 0.12, outlier.shape = NA, alpha = 0.8, linewidth = 0.7) +
      scale_fill_manual(values = COND_COLORS) +
      scale_color_manual(values = COND_COLORS) +
      theme_mous() +
      theme(legend.position = "none") +
      labs(
        title = paste("Pre-stimulus beta power by condition -", subject),
        subtitle = "Mixed-effects model controls for position-in-block and block random effects",
        x = NULL, y = "Pre-stim beta power (a.u.)"
      )
  } else {
    cat("No prestim_beta/condition columns found; Aim 1 figure skipped.\n")
  }
} else {
  cat("No trial-level CSV found; Aim 1 figure skipped.\n")
}
```

# Aim 2 - MEG-fMRI Coupling

## m11 Governance (Primary vs Exploratory)

```{python}
m11 = metrics.get("m11_coupling", {}) if isinstance(metrics.get("m11_coupling"), dict) else {}
rows = []
for key, val in m11.items():
    if not (isinstance(key, str) and key.endswith("_vs_mtg") and isinstance(val, dict)):
        continue
    feature = key.split("_vs_mtg")[0].split("_", 1)[1] if "_" in key else key
    rows.append(
        {
            "contrast": key,
            "feature": feature,
            "role": (m11.get("feature_roles", {}) or {}).get(feature, "exploratory"),
            "r": val.get("r"),
            "p": val.get("p"),
            "p_fdr_bh": val.get("p_fdr_bh"),
        }
    )
if rows:
    pd.DataFrame(rows).sort_values(["role", "contrast"]).style.format(precision=4)
else:
    print("No m11 per-feature coupling entries available.")
```

```{r}
#| code-fold: true

joined_file <- file.path(export_dir, paste0(subject, "_trials_joined.csv"))
if (file.exists(joined_file)) {
  joined <- read_csv(joined_file, show_col_types = FALSE)
  req_cols <- c("mtg_beta", "prestim_beta", "n400m", "dci_trial", "condition", "pos_in_block", "block_id")
  if (all(req_cols %in% names(joined))) {
    joined$condition <- factor(joined$condition)
    fit <- lmer(
      mtg_beta ~ prestim_beta + n400m + dci_trial + condition + pos_in_block + (1 | block_id),
      data = joined
    )
    summary(fit)
  } else {
    missing_cols <- setdiff(req_cols, names(joined))
    knitr::asis_output(
      paste0(
        "Aim 2 model skipped (missing joined columns): ",
        paste(missing_cols, collapse = ", ")
      )
    )
  }
} else {
  knitr::asis_output("Joined MEG-fMRI CSV not available; Aim 2 model skipped.")
}
```

```{r}
#| label: fig-meg-fmri-scatter
#| fig-cap: "MEG pre-stimulus beta power versus left-MTG fMRI beta weight."
#| fig-width: 8
#| fig-height: 5

if (file.exists(joined_file)) {
  joined <- read_csv(joined_file, show_col_types = FALSE)
  req_cols <- c("prestim_beta", "mtg_beta", "condition")
  if (all(req_cols %in% names(joined))) {
    joined$condition <- factor(joined$condition)
    ggplot(joined, aes(x = prestim_beta, y = mtg_beta, color = condition)) +
      geom_point(alpha = 0.45, size = 1.8) +
      geom_smooth(method = "lm", se = TRUE, linewidth = 1, alpha = 0.15) +
      scale_color_manual(values = COND_COLORS, name = "Condition") +
      theme_mous() +
      theme(legend.position = "right") +
      labs(
        title = paste("MEG-fMRI coupling -", subject),
        subtitle = "Pre-stimulus beta power predicts left-MTG BOLD response",
        x = "Pre-stim beta power (MEG, a.u.)",
        y = "Left-MTG beta weight (fMRI)"
      )
  } else {
    cat("No prestim_beta/mtg_beta/condition columns found; Aim 2 figure skipped.\n")
  }
} else {
  cat("No joined trial CSV found; Aim 2 figure skipped.\n")
}
```

# Aim 3 - Traveling Waves

```{r}
#| label: fig-wave-density
#| fig-cap: "Smoothed angular density curves for each condition."
#| fig-width: 6
#| fig-height: 5

if (file.exists(dirs_file)) {
  dirs <- read_csv(dirs_file, show_col_types = FALSE)
  ggplot(dirs, aes(x = direction_rad, color = condition, fill = condition)) +
    geom_density(alpha = 0.2, linewidth = 0.9, bw = 0.3) +
    coord_polar() +
    scale_color_manual(values = COND_COLORS, name = "Condition") +
    scale_fill_manual(values = COND_COLORS, name = "Condition") +
    theme_mous() +
    theme(legend.position = "right") +
    labs(
      title = paste("Wave direction density -", subject),
      subtitle = "Smoothed density overlaid across conditions",
      x = NULL, y = NULL
    )
} else {
  cat("No direction CSV found; Aim 3 density skipped.\n")
}
```

## Source-space Summary (m5)

```{python}
m5_keys = [
    "source_dci_zinnen",
    "source_directions_count",
    "m5_n_labels",
    "m5_n_stcs",
    "m5_error",
    "m5_skipped_reason",
]
m5_labels = {
    "source_dci_zinnen": "Source DCI - ZINNEN",
    "source_directions_count": "Source direction count",
    "m5_n_labels": "ROI labels",
    "m5_n_stcs": "Source estimates (STCs)",
    "m5_error": "Error",
    "m5_skipped_reason": "Skipped reason",
}
m5_rows = [{"Metric": m5_labels.get(k, k), "Value": _fmt(metrics[k])} for k in m5_keys if k in metrics]
if m5_rows:
    pd.DataFrame(m5_rows).style.hide(axis="index")
else:
    print("m5 metrics not present (stage may have been skipped).")
```

# Integrated Aims Narrative

This section ties the three aims into one analysis story: pre-stimulus preparation (Aim 1), multimodal validation in MTG (Aim 2), and traveling-wave boundary conditions (Aim 3). Each subsection reports expected direction, observed summary metrics, and interpretation-ready diagnostics.

## Aim 1 - Pre-stimulus Beta and N400m Link

```{r}
#| label: fig-aim1-panel-c
#| fig-cap: "Aim 1 panel C equivalent: trial-wise pre-stim beta versus N400m with Spearman trend."
#| fig-width: 7.5
#| fig-height: 4.8

if (file.exists(trials_file)) {
  trials <- read_csv(trials_file, show_col_types = FALSE)
  req_cols <- c("prestim_beta", "n400m", "condition")
  if (all(req_cols %in% names(trials))) {
    trials <- trials[is.finite(trials$prestim_beta) & is.finite(trials$n400m), ]
    if (nrow(trials) >= 5) {
      cor_res <- suppressWarnings(cor.test(trials$prestim_beta, trials$n400m, method = "spearman"))
      ann <- paste0("Spearman r = ", sprintf("%.3f", unname(cor_res$estimate)), ", p = ", format.pval(cor_res$p.value, digits = 3))
      ggplot(trials, aes(x = prestim_beta, y = n400m, color = condition)) +
        geom_point(alpha = 0.45, size = 1.7) +
        geom_smooth(method = "lm", se = TRUE, linewidth = 0.9, alpha = 0.2) +
        annotate("label", x = Inf, y = -Inf, hjust = 1.05, vjust = -0.5, label = ann, size = 3.5) +
        scale_color_manual(values = COND_COLORS, name = "Condition") +
        theme_mous() +
        theme(legend.position = "right") +
        labs(
          title = paste("Aim 1: Pre-stim beta vs N400m -", subject),
          subtitle = "More negative N400m is lower on the y-axis",
          x = "Pre-stim beta power (a.u.)",
          y = "N400m amplitude (a.u.)"
        )
    } else {
      cat("Insufficient finite trials for Spearman summary.\n")
    }
  } else {
    cat("Aim 1 correlation skipped (missing prestim_beta/n400m/condition).\n")
  }
} else {
  cat("Trials CSV not available.\n")
}
```

```{python}
aim1_rows = []
for key in (
    "aim1_n400m_zinnen_vs_woorden_t",
    "aim1_prestim_block_effect_p",
    "aim1_prestim_first_trial_effect_p",
    "aim1_prestim_auc",
    "aim1_prestim_auc_guard_reason",
):
    if key in metrics:
        aim1_rows.append({"Metric": key, "Value": _fmt(metrics[key])})
pd.DataFrame(aim1_rows).style.hide(axis="index") if aim1_rows else print("No Aim 1 metrics available.")
```

## Aim 2 - MEG-fMRI Correspondence at Left MTG

```{python}
import numpy as np
import matplotlib.pyplot as plt

lag_coupling = metrics.get("m11_coupling_hrf_lag_sweep", {}) if isinstance(metrics.get("m11_coupling_hrf_lag_sweep"), dict) else {}
if lag_coupling:
    xs, ys = [], []
    for label, row in lag_coupling.items():
        try:
            lag_val = float(str(label).replace("s", ""))
        except Exception:
            continue
        z_row = row.get("zinnen_prestim_beta_vs_mtg", {}) if isinstance(row, dict) else {}
        r = z_row.get("r")
        if r is None:
            continue
        xs.append(lag_val)
        ys.append(float(r))
    if xs:
        order = np.argsort(xs)
        xs = np.asarray(xs)[order]
        ys = np.asarray(ys)[order]
        fig, ax = plt.subplots(figsize=(7.5, 4.0))
        ax.plot(xs, ys, marker="o", linewidth=1.8, color="#2563eb")
        ax.axhline(0, linestyle="--", linewidth=1.0, color="#94a3b8")
        best_i = int(np.nanargmax(np.abs(ys)))
        ax.scatter([xs[best_i]], [ys[best_i]], color="#dc2626", zorder=3)
        ax.set_title("Aim 2 lag optimization (ZINNEN prestim-beta vs MTG)")
        ax.set_xlabel("Assumed hemodynamic lag (s)")
        ax.set_ylabel("Spearman r")
        plt.show()
    else:
        print("Lag sweep metric present but no parseable lag entries.")
else:
    print("No HRF lag sweep coupling entries available.")
```

```{python}
aim2_rows = []
m11 = metrics.get("m11_coupling", {}) if isinstance(metrics.get("m11_coupling"), dict) else {}
for key in ("zinnen_prestim_beta_vs_mtg", "woorden_prestim_beta_vs_mtg"):
    row = m11.get(key, {})
    if isinstance(row, dict):
        aim2_rows.append({"Contrast": key, "r": row.get("r"), "p": row.get("p"), "p_fdr_bh": row.get("p_fdr_bh")})
if aim2_rows:
    pd.DataFrame(aim2_rows).style.format(precision=4)
else:
    print("No condition-specific coupling rows available in m11_coupling.")
```

## Aim 3 - Traveling-Wave Directionality

```{python}
import numpy as np

m12_dir = export_dir.parent.parent / "m12_wave_validation"
null_path = m12_dir / f"sub-{subject}_null_dci.npy"
empirical = float(metrics.get("dci_zinnen", np.nan))
if null_path.exists():
    null_dci = np.load(null_path)
    fig, ax = plt.subplots(figsize=(7.2, 4.0))
    ax.hist(null_dci, bins=30, color="#cbd5e1", edgecolor="white")
    if np.isfinite(empirical):
        ax.axvline(empirical, color="#dc2626", linewidth=2, linestyle="--")
    ax.set_title("Aim 3 two-dipole null comparison (ZINNEN DCI)")
    ax.set_xlabel("Null-model DCI")
    ax.set_ylabel("Count")
    plt.show()
else:
    print("Two-dipole null DCI artifact not found; histogram skipped.")
```

```{python}
aim3_rows = []
for key in (
    "dci_zinnen",
    "dci_rest",
    "p_task_vs_rest",
    "aim3_two_dipole_z",
    "aim3_two_dipole_z_threshold",
    "aim3_two_dipole_passes_threshold",
):
    if key in metrics:
        aim3_rows.append({"Metric": key, "Value": _fmt(metrics[key])})
pd.DataFrame(aim3_rows).style.hide(axis="index") if aim3_rows else print("No Aim 3 metrics available.")
```
